A Recipe for Scavenging - the natural history of a behaviour

### 2016-04-25 natural history of scavenging.tex
trying to get this pull request to go through with pointless edit
reordered sections, fixed up detection and handling time

\begin{abstract} 
% Here, we chart the natural history of scavenging. 
Despite its prevalence, scavenging is a difficult behaviour to observe in modern day carnivores and impossible to study directly in extinct species. 
Yet, there are certain intrinsic and environmental features of a species that push it towards a scavenging lifestyle. 
These can be thought of as some of the principal parameters in optimal foraging theory namely, encounter rate, handling time and prey availability. 
% Chief among these are low-cost locomotion, high detection distances, effective carcass processing and a carrion-producing habitat. 
We use these components to highlight the morphologies and environments that would have been conducive to scavenging over geological time by focusing on the dominant vertebrate groups of the land, sea and air. 
The result is a document on the natural history of scavenging, the first to our knowledge. 
Our idea of a scale of scavenging can be applied to any species at any time to judge the importance of this behaviour in its diet. 
\end{abstract}

@@

Historically, scavengers have not been viewed as the most charismatic of animals.
This may go some way to explaining the gap in our knowledge of the prevalence of this behaviour \citep{devault2003scavenging}.
Professor Sanborn Tenney writing in 1877 for The American Naturalist had this to say about one well known group, ``prominent among the mammalian scavengers are the hyenas, the ugliest in their general appearance of all the flesh eaters."
He contrasts these with ``nobler kinds" of carnivores such as lions and tigers \citep{tenney1877few}.
Even aside from our own subjective biases, scavenging is a difficult behaviour to detect after the fact.
Without catching a carnivore in the act of killing we are left to infer how the prey was killed.
Some simple heuristics can inform us, for instance, in cases where the prey item was simply too large to have been killed by the ostensible predator \citep{pobiner2008paleoecological}.
But clearly, a scavenger does not only feed on animals too big for it to have hunted.
The obvious lack of direct behavioural data compounds the difficulty of discerning scavenging from predation among extinct forms.
Indeed, a single species of dinosaur notwithstanding \citep{carbone2011intra}, a synthesis describing the natural history of scavengers is absent from the literature.
%TG: I don't understand the dinosaur part of this sentence. Might be my non-native english but I thought it's worth noting for other non-english folk out there (even though I know they aren't many!).
With research on scavenging on the rise \citep{manga2006vulture} we are now beginning to realise the extent of this behaviour such that, ``in some ecosystems, vertebrates have been documented to assimilate as much as 90\% of the available carrion" \citep{beasley2015vertebrates}.
This has profound implications for the trophic ecology of these systems and particularly our models of them.
Even Tenney’s noble big cats are now known to take in a significant portion of carrion in their diet where some lion populations acquire over 50\% of their meat from carcasses \citep{jones2015african}.
While recognising the difficulty in directly observing scavenging, it is possible to turn to other methods in order to discern the most suitable morphologies, physiologies and environments for a scavenging lifestyle to prosper.
Here we chart the natural history of scavenging by assessing the potential for the behaviour in dominant vertebrate groups given their ecology and functional traits.
%It is our aim in this review to employ these methods to gain an understanding of scavengers past and present.

@@

\subsection*{The Challenges of Scavenging} %TG: challenges or difficulty? up to you.
The chief hurdle to scavenging is finding a resource that is often difficult to predict in space and time.
Through chance alone many species will avail of some opportunistic scavenging. %TG: Two times "opportunities", replaced by "chances"
However, species that rely on scavenging to sustain substantial portions of their diets must encounter a sufficient amount of carrion in order to meet their energetic demands.
%Once found, the scavenger must be able to out-compete any potential competitors and process a product under the processes of decay from toxin producing and disease causing microorganisms \citep{ruxton2014fruit}. %TG: make the following sentence more sense?
Once found, the scavenger must be able to out-compete any potential  competitors and process the, typically decaying, carcass replete with microorganism derived toxins  \citep{ruxton2014fruit}. Both these characteristics can be assumed to be under evolutionary selection pressures for traits that increase carrion discovery and monopoly.
Finally, the potential for scavenging will also depend on the density, size, and quality of carcases produced, all of which are affected by complex ecosystem dynamics and are outside the selection pressures on the scavenger.
Each of these facets are essentially the backbone of fundamental ecological theory and are the key parameters defined in functional response curves, namely encounter rate, handling time and prey availability \citep{jeschke2002predator}. %TG: Jaysus, this sentence doesn't read well! "Functional response curves of optimal foraging theory" can probably be phrased in a nicer/easier way...
%With this framework we compare the ability of scavenging across vertebrates. % focusing on each of these facets and their relationship with scavenging.
By considering scavenging in this context of optimal foraging we can identify the prerequisite attributes and processes required for the behaviour. 
This has enabled us to propose a scale of scavenging whereupon we can place any vertebrate species, past or present, and assess the importance of carrion in its diet.

@@

%KH: need paragraph for how do we know something is (was) a scavenger
\section*{Encounter Rate}
All foraging processes depend on the encounter rate between consumer and resource. 
In the simplest case, this rate can be thought of in terms of a gas diffusion model where the movement of two agents (i.e. predator and prey) depends only on as they can actively detect resources through their sensory abilities. %TG: I like the gas diffusion analogy but it might be adding too much text and be not essential. I've added some precisions though.
%KH I'v shortened it some more.
As carcasses are stationary, the relative speed between a scavenger and carrion is only dependent on the movement of the scavenger.
%KH There are too many definite statements throughout from my liking. For example, not all carcasses are stationary in fact the sinking of carcasses in marine systems is what drives whale fall fauna as they essentially let the carcass come to them. I know it might make the text more clunky but I would play it safe and "As carcasses are usually stationary" etc.
As such, scavenging potential is strongly affected by search rates, which are determined by both species' physiology and what type of environment the scavenger is foraging in.
%KH is including dimensionality making this too complex as the only search environment that is 3d is for semi-buoyant marine carcasses. Also the dimensionality talked about in the vision section is different to this dimensionality (the Dim of the search space for both terrestrial and aerial are both 2D). Might be a bit confusing so I changed it to just the physiology and environmental

\subsection*{Locomotion}
Because of the inherent unpredictability of carrion, scavenging heavily depends on the ability to efficiently move over large areas than predation.
%KH changed "more" to "heavily"
This generally requires an efficient transfer of metabolic energy into movement which relies on both physiology (i.e. metabolism) and the medium of the environment in which the animal is moving (i.e. aerial, aquatic or terrestrial).
Perhaps the most efficient form of locomotion in vertebrates is, paradoxically, found in flying species. 
Despite the energetic costs of flight, the only known vertebrate obligate scavengers are the old and the new world vultures. 
And, although powered flight is energetically expensive, species like vultures can exploit air currents using their large wingspans which allows them to soar at a cost of only twice their metabolic rate \citep{hedenstrom1993migration,spivey2014analysing}.
%KH This and seems out of place a bit, maybe replace with "While powered flight..."
By depending on thermal air flows these species can forage over vast ranges \citep{spiegel2013factors}. 
An analogous mode of locomotion is also exploited by seabirds, who use strong ocean winds to search large areas of the oceans \citep{norberg2012vertebrate,thaxter2012seabird}. 
While many species of seabird are likely primarily predators, it seems that albatrosses, who can range many hundreds of kilometres, take a substantial amount of carrion in their diet \citep{croxall1994dead}. 
This is typically in the form of squid carcases, which float on the surface, allowing the birds to readily pluck their remains out of the water \citep{croxall1994dead}.

The groups from which these modern soaring birds arose, appear during the Palaeocene \citep[66 - 56 Million years ago (Mya); ][]{Jetz2012, Jarvis2014} and Cretaceous \citep{chiappe2006early} respectively.  % KH and are likely to have being soaring fliers as demonstrated by their large wingspans (Is this true?). (Is there any other form of soaring?).  AK I don't think this is true from what I've seen
%KH OK I changed it to these modern soaring birds to emphasize we are still on the tred of low cost flying. However if you don't think its clear enough what this group is feel free to change it back.
However, soaring flight is likely to be far older than this with avian flight originating in the Late Jurassic (163.5-145 Mya) and vertebrate flight in the Late Triassic (235-201.3 Mya) coincident with the pterosaurs. 
Indeed, scavenging among pterosaurs has been hypothesised many times before \citep{witton2008reappraisal}. 
Certain groups of these animals could reach enormous sizes \citep[e.g. Azhdarchids with wingspans of 11 metres; ][]{witton2010size} and, notably, appear to have engaged in soaring flight \citep{witton2010size}.
It seems probable that as least some of these extinct species used soaring as a means for scavenging \citep{witton2013pterosaurs}. 
%KH added "at least some of these" as it was not clear what species used soaring as a means for scavenging.
While soaring is perhaps the only viable means of locomotion that allows for an obligate, scavenging life-style \citep{ruxton2004obligate}, powered flight is still an efficient means of locomotion. 
Certainly, avian flight is cheaper than either walking or running \citep{tucker1975energetic}.
%TG: I'd rather put the following part as an appendix or foot note. Reader can take our (Tucker's) word for it and then check the math to see that even with maintenance and blablabla flying wins.
%Even taking account of maintenance costs, this still bears out, where total cost of movement (J kg\textsuperscript{-1} m\textsuperscript{-1}) scales according to 5.2 $\times$ body mass (kg) \textsuperscript{-0.23} for fliers and 10.7 $\times$ body mass (kg) \textsuperscript{-0.32} for runners \citep{williams1999evolution}.  
%TG: This also allows to link directly to the next argument in the same paragraph: facultative active flying scavengers.
%KHI agree avoid as much maths as possible (If only someone would propose some sort of drop down thing for maths in papers.)
We know that many extant birds exist as facultative scavengers because storks, raptors and corvids all take substantial quantities of carrion in their diet \citep{kendall2013alternative}. 
Similarly we would expect that extinct species would also scavenge in a similar fashion depending on the efficiency of their flight. 
For example, early birds such as \textit{Archaeopteryx} are predicted to have been poor, relatively inefficient fliers \citep{nudds2010narrow} and so ill-suited to finding carrion. 
%Similarly something about the efficiency of pterosaurs flight. TG: Was this a comment? By the way, about pterosaurs (though it might not be necessary to mention it), most of them where tiny little feckers like Archaeopteryx or even smaller.

The importance of efficient flying over large areas may explain the lack of scavenging behaviour in bats as they are generally nocturnal, a time when they would receive no aid from convective air currents \citep{norberg2012vertebrate}. 
%smaller and less efficient in comparison to larger birds (speculating here but I wouldnt be surprised), %AK Don't think it is the case that bats are less efficient than birds in terms of flight 
That said, \textit{Necromantis} (``death-eater''), a large bat from the middle to late Eocene (56-33.9 Mya) had a robust cranio-mandibular morphology, and is a likely candidate for scavenging behaviour \citep{Weithofer_Necromantis_1887,Hand_Necromantis_2012}.% amongst the x amount of known bat species \citep{Weithofer_Necromantis_1887,Hand_Necromantis_2012}. %TG: we can cut this number of bat species to avoid getting into stamp collecting debate. + there are effectively a two shit load units of them which makes the Necromantis example not really convincing.

Similar to aerial species, aquatic scavengers have a locomotory benefit because water is a medium that is conducive to low-cost movement \citep{tucker1975energetic,williams1999evolution}.
%TG: Same as comment before, maybe put the following as a footnote or as a supplementary.
% In fact, the total cost of movement (again in J kg\textsuperscript{-1} m\textsuperscript{-1}) in salmonids is lower than either running or flying where it scales according to 2.15 $\times$ body mass (kg) \textsuperscript{-0.25} \citep{williams1999evolution} with only \textit{soaring} flight likely to surpass it. 
%KH(I assume the reference is for powered flight, again this needs to be more precise). 
%KH again I think its good  to keep away from the maths 
This has led some researchers to argue for the likelihood of an obligate scavenging fish \citep{ruxton2004energetic,ruxton2005searching}.

Sharks are one likely candidate for general scavenging behaviors as their locomotion, which depends on large pectoral fins in order to maintain lift as they swim, resemble large soaring fliers.
%changed this sentence up a bit 
Many shark species have large foraging ranges \citep[e.g. the great white sharks \textit{Carcharodon carcharias};][]{bruce2006movements} and it seems reasonable that they would use oceanographic currents to further reduce movement costs \citep{ruxton2004energetic}. %and are able to maintain a type of endothermy from thier movements \citep{bernal2015sharks}. %TG: I guess the following is a comment: (I know the Greenland shark is the slowest swimmer). 
In fact, facultative scavenging is seen in many selachian groups, including species of extant sharks like white sharks \citep[known to feed on whale carcasses;][]{fallows2013white}, Greenland sharks \citep[feeding on seals;][]{watanabe2012slowest}, and sixgill sharks \citep{anderson2016impact}. 
The former which grow up to 6 metres long, can be sustained by 30 kg of whale blubber for over six weeks \citep{carey1982temperature}. 
%A personal thing but I hate the use of former or latter as I have to break my reading and go back and see which species that is. I did'nt change it but I would go with naming the species there and then for clarity.
There is evidence too of scavenging in extinct species, where shark teeth have been found in the remains of dinosaurs, mosasaurs and Pliocene mysticete whales \citep[5.3-3.6 Mya; ][]{schwimmer1997scavenging,ehret2009caught}.

Interestingly, style of swimming in fish does not significantly affect the cost of movement \citep{williams1999evolution}, hence it is likely that aquatic species with large ranges may also encounter scavenging opportunities. It may be that jaw morphology is more likely to define which species can avail of carcasses.
%KH I moved this down here as it kinda throws our shark soaring thing out the window (which I like as a linking function). However I think this kinda solves our problems in the complexity of swimming by saying apart from large ranges ability to eat is probably more important.
%KH also is fish an ok term to use, will people get pissy or anything like with reptiles?

% KH endotermy - why it doesn't seem to work
We might expect then that by combining an aquatic environment and an endothermic metabolism marine mammals would especially prosper as scavengers, especially as this group has perhaps the largest home range of any animals (whales and shit).
Fossil pinnipeds and cetaceans from 60 Mya %TG: that early???
have transitional features indicative of their trajectory to fully aquatic species \citep{williams1999evolution}.  
But despite their movement away from land their energetic savings were negligible because the \textit{total} cost incurred by a swimming marine mammal is high \citep{williams1999evolution}. 
Indeed, the total energetic cost is similar to an equivalent terrestrial or aerial mammal \citep{williams1999evolution}.
This underscores the trade offs between the benefits of endothermy in terms of activity periods and the costs of maintaining such an energetically costly system. 
That said, aquatic endotherms have and do scavenge. 
For instance, early whales such as \textit{Basilosaurus} (38-36.5 Mya) seem to have fit into the same niche as killer whales (\textit{Orcinus orca}) and we have some evidence for scavenging in both \citep{fahlke2012bite,Whitehead415}.

@@

%\cite{williams1999evolution} notes, ``free-ranging animals must contend with the total energetic expenditure associated with supporting basic biological functions as well as with moving the body and appendages through the environment."

@@

%TG: Hmm I get the point of gas exchange efficiency but (1) it should be more developed if left like that (i.e. what the flip is meant by gas efficient exchanges) and (2) that works for mammals, but in terms of terrestrial that ignores birds, squamates, crocodiles and amphibians.

@@

The early transition from a sprawling gait, such as seen in Tiktaalik, to the more erect posture of Synapsids and later dinosaurs and mammals, has often been supposed as conferring a huge advantage to the latter.
%KH Even after changing it I dont like this transition type narrative.
%KH also needs to be more clear on what the advantage to scavenging is for these groups.
%The purported advantages include benefits in terms of speed, efficiency, muscle effort and manoeuvrability \citep{sullivan2015posture}. 
%none of these advantages refer very clearly to scavenging
Despite being intuitive, \cite{sullivan2015posture} states most of the hypotheses in favour of this idea remain to be tested in the context of archosaur evolution. 
%KH I'm always wary of deeming something as intuitive,

@@

%This is exacerbated by their sprawling gait which results in the phenomenon known as Carrier's constraint such that the animal can't move and breathe at the same time because the lateral movements impedes its lungs \citep{carrier1987evolution}. 
% These constraints are likely to mean that many ectotermic terrestrial species, which cannot maintain an upright posture would depend on a relatively low level of scavenging (I'm not sure this is true though). 
%This would also have been true of extinct species with the same physiology. 
%TG: I guess this is a comment too: Snakes are a bit different are more effecent at moving I think - lots of scavanging. 
%Both these groups are more likely to depend on low energetic costs in order to compensate for low encounter rates due to their inability to move. 
Whatever the case, it is with the evolution of endothermy in the therapsid-mammal lineage \citep{clarke2010temperature} that terrestrial vertebrates would have gained the ability to range more widely, a vital component in seeking out carrion. 
Although the earliest evidence of vertebrate scavenging comes from the Permian (298.9 - 252.17 Mya) where a temnospondyl amphibian fed on the carcass of \textit{Varanops}, a predatory synapsid of the time \citep{reisz2006articulated}. 
%TG: I do like the way all our paragraphs end up with a "But wait, there's an exception!"

Modern endothermic mammals can sustain longer periods of energetically expensive activity \citep{bennett1979endothermy} resulting in larger foraging ranges. 
%TG: Same comment as for the other math, I'd rather put that as a footnote or in the appendix (the whole yoke). Again, it seems a bit disproportionate that for some example, we just dumb a line (or even an example in brackets) and that for some others, we explain the example in details.
% To quantify this effect with a simple example we can turn to some allometric relationships relating sustainable travelling speed to body mass \citep{ruxton2004obligate}.
% If we insert these into a foraging radius model \citep{Enstipp2006Energetics} for a 12 hour foraging day it shows that while a 10 kg reptile can range 6.5 km, an equally sized mammal can range nearly 33 km (See appendix). 
% For a foraging scavenger, this ability translates into a greater area searched for food.
Today, terrestrial scavenging in the mammals is probably best known in an African context where hyenas, jackals and lions all take sizeable proportions of carrion in their diet.
In the spotted hyena (\textit{Crocuta crocuta}), striped hyena (\textit{Hyaena hyaena}) and brown hyena (\textit{Hyaena brunnea}) it can be over 90\% \citep{jones2015african}.
And although no contemporary terrestrial vertebrate exists as an obligate scavenger, most, if not all, are facultative to some extent \citep{beasley2015vertebrates}.
The particular reliance of hyenas on carrion means we can use them as examples of efficient terrestrial scavengers to compare with other forms. 
In terms of locomotion, they employ a characteristic ``rocking horse gait"  which allows them to cover great distances efficiently, loping at 10 km/hr \citep{mills1989comparative,jones2015african}. 
Such long-distance travel is apparent in African wild dogs (\textit{Lycaon pictus}) and many other canids \citep{pennycuick1995radius,janis2014forelimb}. 
In contrast, big cats like leopards (\textit{Panthera pardus}) rely on ambush \citep{pennycuick1995radius}. 
This allows us to make a broad distinction between the ambush strategies of cats-like forms and the pursuit/ pounce strategies of more dog-like forms%TG: I've added the "-like" to underline that we're not talking about the phylogeny here (i.e. not Felidae (incl. hyena) vs Canidae)
, the latter being more suited to scavenging \citep{janis2014forelimb}. 
These insights allow us to compare extant terrestrial species to their prehistoric forebears given the dominance of mammalian carnviores since the Eocene (56-33.9 Mya) where the order split into the Caniforma and Feliforma \citep{van1987skeletal}.
To take one example, \cite{anyonge1996locomotor} found that \textit{Nimravides}, a genus of sabretooth cat from the Miocence (10.3 to 5.3 Mya), were likely to have been ambush predators which would argue against them taking a lot of carrion. 
%The order Carnivora saw its origins in the Middle Eocene  %TG: a phylo side note: the correct sentence would be the "The stem Carnivora saw its origins ... when it split into Caniforma and Feliforma." But I think we can leave it out because, god, it sounds pedantic!
% So we can trace efficient terrestrial movement by carnivores from this point on.

While we can use modern analogues of quadrupedal species, the absence of scavenging in bipedal animals makes this more difficult. However, as a groups bipedal locomotion may have been a particularly useful strategy for  searching large areas. For instance, \cite{alexander2004bipedal} shows that, in the case of humans, we are more economical than predicted while walking according to predicted costs of terrestrial movement. In fact the ability of humans to engage in hunting strategies that involve traversing long ranges, such as persistence hunting, has led to several researches highlighting the possible role of scavenging in early hominids. Indeed the low energetics of human walking at least allows for a substantial amount of scavenging in our hominid ancestors (Graemes papers on scavenging).
%%KH I know this is excluding other hominids so feel free to write a more inclusive line.

>>>>

\begin{abstract} 
Despite its prevalence, scavenging is a difficult behaviour to observe in modern day carnivores and impossible to study directly in extinct species. 
Yet, there are certain intrinsic and environmental features of a species that push it towards a scavenging lifestyle. 
These can be thought of as some of the principal parameters in optimal foraging theory namely, encounter rate, handling time and prey availability. 
We use these components to highlight the morphologies and environments that would have been conducive to scavenging over geological time by focusing on the dominant vertebrate groups of the land, sea and air. 
The result is a document on the natural history of scavenging, the first to our knowledge. 
Our idea of a scale of scavenging can be applied to any species at any time to judge the importance of this behaviour in its diet. 
\end{abstract}

@@

Historically, scavengers have not been viewed as the most charismatic of animals.
This may go some way to explaining the gap in our knowledge of the prevalence of this behaviour \citep{devault2003scavenging}.
Professor Sanborn Tenney writing in 1877 for The American Naturalist had this to say about one well known group, ``prominent among the mammalian scavengers are the hyenas, the ugliest in their general appearance of all the flesh eaters."
He contrasts these with ``nobler kinds" of carnivores such as lions and tigers \citep{tenney1877few}.
Even aside from our own subjective biases, scavenging is a difficult behaviour to detect after the fact.
Without catching a carnivore in the act of killing we are left to infer how the prey was killed.
Some simple heuristics can inform us, for instance, in cases where the prey item was simply too large to have been killed by the ostensible predator \citep{pobiner2008paleoecological}.
But clearly, a scavenger does not only feed on animals too big for it to have hunted.
The obvious lack of direct behavioural data compounds the difficulty of discerning scavenging from predation among extinct forms.
Indeed, a single species of dinosaur notwithstanding \citep{carbone2011intra}, a synthesis describing the natural history of scavengers is absent from the literature.
With research on scavenging on the rise \citep{manga2006vulture} we are now beginning to realise the extent of this behaviour such that, ``in some ecosystems, vertebrates have been documented to assimilate as much as 90\% of the available carrion" \citep{beasley2015vertebrates}.
This has profound implications for the trophic ecology of these systems and particularly our models of them.
Even Tenney’s noble big cats are now known to take in a significant portion of carrion in their diet where some lion populations acquire over 50\% of their meat from carcasses \citep{jones2015african}.
While recognising the difficulty in directly observing scavenging, it is possible to turn to other methods in order to discern the most suitable morphologies, physiologies and environments for a scavenging lifestyle to prosper.
Here we chart the natural history of scavenging by assessing the potential for the behaviour in dominant vertebrate groups given their ecology and functional traits.
%It is our aim in this review to employ these methods to gain an understanding of scavengers past and present.

@@

\subsection*{The Challenges of Scavenging} 
The chief hurdle to scavenging is finding a resource that is often difficult to predict in space and time.
Through chance alone many species will avail of some opportunistic scavenging. 
However, species that rely on scavenging to sustain substantial portions of their diets must encounter a sufficient amount of carrion in order to meet their energetic demands.
Once found, the scavenger must be able to out-compete any potential competitors and process the, typically decaying, carcass replete with microorganism derived toxins  \citep{ruxton2014fruit}. 
Both of these characteristics can be assumed to be under evolutionary selection pressures for traits that increase carrion discovery and monopoly.
Finally, the potential for scavenging will also depend on the density, size, and quality of carcases produced, all of which are affected by complex ecosystem dynamics and are outside the selection pressures on the scavenger.
Each of these facets are essentially the backbone of fundamental ecological theory and are the key parameters defined in functional response curves, namely encounter rate, handling time and prey availability \citep{jeschke2002predator}.  
By considering scavenging in this context of optimal foraging we can identify the prerequisite attributes and processes required for the behaviour. 
This has enabled us to propose a scale of scavenging whereupon we can place any vertebrate species, past or present, and assess the importance of carrion in its diet.

@@

%KH: need paragraph for how do we know something is (was) a scavenger
\section*{Encounter Rate}
All foraging processes depend on the encounter rate between consumer and resource. 
Locomotory speed, foraging time and detection radius all increase the encounter rate between a scavenger and the carcasses it is searching for. 
We would thus expect selection pressures to act on the various traits that govern these parameters.  
% However, encounter rate is also determined by the productivity of the environment, something which the selective forces acting on the scavenging have no control over.

%As such, scavenging potential is strongly affected by search rates, which are determined by anatomy, physiology and behaviour which are all under selective pressure, as well as the type of environment the scavenger is foraging in.

\subsection*{Locomotion}
Because of the inherent unpredictability of carrion, scavenging depends more on the ability to efficiently move over large areas than predation.
This generally requires an efficient transfer of metabolic energy into movement which relies on the animal's anatomy and physiology as well as the medium of the environment in which the animal is moving (i.e. aerial, aquatic or terrestrial).
Perhaps the most efficient form of locomotion in vertebrates is, paradoxically, found in flying species. 
Despite the energetic costs of flight, the only known vertebrate obligate scavengers are the old and the new world vultures. 
While powered flight is energetically expensive, species like vultures have evolved behavioural and anatomical features to exploit air currents using their large wingspans, allowing them to soar at a cost of only twice their metabolic rate \citep{hedenstrom1993migration,spivey2014analysing}.
By depending on thermal air flows these species can forage over vast ranges \citep{spiegel2013factors}. 
An analogous mode of locomotion is also exploited by seabirds, who use strong ocean winds to search large areas of the oceans \citep{norberg2012vertebrate,thaxter2012seabird}. 
While many species of seabird are likely primarily predators, it seems that albatrosses, who can range many hundreds of kilometres, take a substantial amount of carrion in their diet \citep{croxall1994dead}. 
This is typically in the form of squid carcases, which float on the surface, allowing the birds to readily pluck their remains out of the water \citep{croxall1994dead}.

The groups from which these modern soaring birds arose, appear during the Palaeocene \citep[66 - 56 Million years ago (Mya); ][]{Jetz2012, Jarvis2014} and Cretaceous \citep{chiappe2006early} respectively. 
However, soaring flight is likely to be far older than this with avian flight originating in the Late Jurassic (163.5-145 Mya) and vertebrate flight in the Late Triassic (235-201.3 Mya) coincident with the pterosaurs. 
Indeed, scavenging among pterosaurs has been hypothesised many times before \citep{witton2008reappraisal}. 
Certain groups of these animals could reach enormous sizes \citep[e.g. Azhdarchids with wingspans of 11 metres; ][]{witton2010size} and, notably, appear to have engaged in soaring flight \citep{witton2010size}.
It seems probable that as least some of these extinct species used soaring as a means for scavenging \citep{witton2013pterosaurs}. 
%KH added "at least some of these" as it was not clear what species used soaring as a means for scavenging.
While soaring is perhaps the only viable means of locomotion that allows for an obligate, scavenging life-style \citep{ruxton2004obligate}, powered flight is still an efficient means of locomotion. 
Certainly, avian flight is cheaper than either walking or running \citep{tucker1975energetic}.

We know that many extant birds exist as facultative scavengers because storks, raptors and corvids all take substantial quantities of carrion in their diet \citep{kendall2013alternative}. 
Similarly we would expect that extinct species would also scavenge in a similar fashion depending on the efficiency of their flight. 
For example, early birds such as \textit{Archaeopteryx} are predicted to have been poor, relatively inefficient fliers \citep{nudds2010narrow} and so ill-suited to finding carrion. 
The importance of efficient flying over large areas may explain the lack of scavenging behaviour in bats as they are generally nocturnal, a time when they would receive no aid from convective air currents \citep{norberg2012vertebrate}.

Similar to aerial species, aquatic scavengers have a locomotory benefit because water is a medium that is conducive to low-cost movement \citep{tucker1975energetic,williams1999evolution}.
This has led some researchers to argue for the likelihood of an obligate scavenging fish \citep{ruxton2004energetic,ruxton2005searching}.

Sharks are one likely candidate for general scavenging behaviors as their locomotion, which depends on large pectoral fins to generate lift as they swim, resembles that of the large soaring fliers.
%changed this sentence up a bit 
Many shark species have large foraging ranges \citep[e.g. the great white sharks \textit{Carcharodon carcharias};][]{bruce2006movements} and it seems reasonable that they would use oceanographic currents to further reduce movement costs \citep{ruxton2004energetic}. %and are able to maintain a type of endothermy from thier movements \citep{bernal2015sharks}. %TG: I guess the following is a comment: (I know the Greenland shark is the slowest swimmer). 
In fact, facultative scavenging is seen in many selachian groups, including species of extant sharks like white sharks \citep[known to feed on whale carcasses;][]{fallows2013white}, Greenland sharks \citep[feeding on seals;][]{watanabe2012slowest}, and sixgill sharks \citep{anderson2016impact}. 
There is evidence too of scavenging in extinct species, where shark teeth have been found in the remains of dinosaurs, mosasaurs and Pliocene mysticete whales \citep[5.3-3.6 Mya; ][]{schwimmer1997scavenging,ehret2009caught}.

Interestingly, style of swimming in fish does not significantly affect the cost of movement \citep{williams1999evolution}, hence it is likely that many aquatic species with large ranges will encounter scavenging opportunities. 
It is perhaps jaw morphology that is more likely to define which species can avail of carcasses.
%KH I moved this down here as it kinda throws our shark soaring thing out the window (which I like as a linking function). However I think this kinda solves our problems in the complexity of swimming by saying apart from large ranges ability to eat is probably more important.
%KH also is fish an ok term to use, will people get pissy or anything like with reptiles?

% KH endotermy - why it doesn't seem to work
We might expect then that by combining an aquatic environment and an endothermic metabolism marine mammals would especially prosper as scavengers. %, especially as this group has perhaps the largest home range of any animals (whales and shit).
Fossil pinnipeds and cetaceans from 60 Mya have transitional features indicative of their evolutionary trajectory to fully aquatic species \citep{williams1999evolution}.  
But despite their movement away from land their energetic savings were negligible because the \textit{total} cost incurred by a swimming marine mammal is high \citep{williams1999evolution}. 
Indeed, the total energetic cost is similar to an equivalent terrestrial or aerial mammal \citep{williams1999evolution}.
This underscores the trade offs between the benefits of endothermy in terms of activity periods and the costs of maintaining such an energetically expensive system. 
That said, aquatic endotherms have and do scavenge. 
For instance, early whales such as \textit{Basilosaurus} (38-36.5 Mya) seem to have fit into the same niche as killer whales (\textit{Orcinus orca}) and we have some evidence for scavenging in both \citep{fahlke2012bite,Whitehead415}.

@@

@@

@@

The early transition from a sprawling gait, such as seen in \textit{Tiktaalik}, to the more erect posture of Synapsids and later dinosaurs and mammals, has often been supposed as conferring a huge advantage to the latter.
%KH Even after changing it I dont like this transition type narrative.
%KH also needs to be more clear on what the advantage to scavenging is for these groups.
%The purported advantages include benefits in terms of speed, efficiency, muscle effort and manoeuvrability \citep{sullivan2015posture}. 
%none of these advantages refer very clearly to scavenging
Despite being intuitive, \cite{sullivan2015posture} states most of the hypotheses in favour of this idea remain to be tested in the context of archosaur evolution. 
%KH I'm always wary of deeming something as intuitive,

@@

%This is exacerbated by their sprawling gait which results in the phenomenon known as Carrier's constraint such that the animal can't move and breathe at the same time because the lateral movements impedes its lungs \citep{carrier1987evolution}. 
% These constraints are likely to mean that many ectotermic terrestrial species, which cannot maintain an upright posture would depend on a relatively low level of scavenging (I'm not sure this is true though). 
%This would also have been true of extinct species with the same physiology. 
%TG: I guess this is a comment too: Snakes are a bit different are more effecent at moving I think - lots of scavanging. 
%Both these groups are more likely to depend on low energetic costs in order to compensate for low encounter rates due to their inability to move. 
Whatever the case, it is with the evolution of endothermy in the therapsid-mammal lineage \citep{clarke2010temperature} that terrestrial vertebrates would have gained the ability to range more widely, a vital component in seeking out carrion. 
Although the earliest evidence of vertebrate scavenging comes from the Permian (298.9 - 252.17 Mya) where a temnospondyl amphibian fed on the carcass of \textit{Varanops}, a predatory synapsid of the time \citep{reisz2006articulated}.

Modern endothermic mammals can sustain longer periods of energetically expensive activity \citep{bennett1979endothermy} resulting in larger foraging ranges. 
Today, terrestrial scavenging in the mammals is probably best known in an African context where hyenas, jackals and lions all take sizeable proportions of carrion in their diet.
In the spotted hyena (\textit{Crocuta crocuta}), striped hyena (\textit{Hyaena hyaena}) and brown hyena (\textit{Hyaena brunnea}) it can be over 90\% \citep{jones2015african}.
And although no contemporary terrestrial vertebrate exists as an obligate scavenger, most, if not all, are facultative to some extent \citep{beasley2015vertebrates}.
The particular reliance of hyenas on carrion means we can use them as examples of efficient terrestrial scavengers to compare with other forms. 
In terms of locomotion, they employ a characteristic ``rocking horse gait"  which allows them to cover great distances efficiently, loping at 10 km/hr \citep{mills1989comparative,jones2015african}. 
Such long-distance travel is apparent in African wild dogs (\textit{Lycaon pictus}) and many other canids \citep{pennycuick1995radius,janis2014forelimb}. 
In contrast, big cats like leopards (\textit{Panthera pardus}) rely on ambush \citep{pennycuick1995radius}. 
This allows us to make a broad distinction between the ambush strategies of cat-like forms and the pursuit/ pounce strategies of more dog-like forms, the latter being more suited to scavenging \citep{janis2014forelimb}. 
These insights allow us to compare extant terrestrial species to their prehistoric forebears given the dominance of mammalian carnviores since the Eocene (56-33.9 Mya) where the order split into the Caniforma and Feliforma \citep{van1987skeletal}.
To take one example, \cite{anyonge1996locomotor} found that \textit{Nimravides}, a genus of sabretooth cat from the Miocence (10.3 to 5.3 Mya), were likely to have been ambush predators which would argue against them taking a lot of carrion.

Of course, terrestrial animals can also move bipedally. 
Although the evolution of bipedal  movement was significant in that it freed up the forelimbs for other purposes (e.g. climbing, tool-use, wing development etc.) it does not differ radically in cost from quadrupedal locomotion (\cite{williams1999evolution}, and references therein). 
For instance, \cite{alexander2004bipedal} shows  that, in the case of humans, we are more economical than predicted while walking and less so while running according to predicted costs of terrestrial movement.
Our locomotory efficiency has fed into the question of where our ancestors placed on the hunter-scavenger axis during the Plio-Pleistocene, which has been a matter of debate for years \citep{dominguez2002hunting}.
\cite{ruxton2013endurance} added to this debate with his argument that long distance endurance running, often used by humans for hunting, was not an important feature of hominin scavenging.
He arrives at this conclusion on the basis of the high cost of running coupled with the fact that dangerous competitors will tend to arrive beforehand.

% While we can use modern analogues of quadrupedal species, the lack of scavenging in bipeds makes this more difficult. However, bipedal locomotion may have been a particularly useful strategy for searching large areas. For instance, \cite{alexander2004bipedal} shows that, in the case of humans, we are more economical than predicted while walking according to predicted costs of terrestrial movement. In fact the ability of humans to engage in hunting strategies that involve traversing long distances, such as persistence hunting, has led to several researches highlighting the possible role of scavenging in early hominids. Indeed the low energetics of human walking at least allows for a substantial amount of scavenging in our hominid ancestors (Graemes papers on scavenging).
%%KH I know this is excluding other hominids so feel free to write a more inclusive line.
